---
title: "MTH404 R Project"
author: "Mustafif Khan"
output: pdf_document
---

# Loading Required Packages
Before the dataset can be analyzed, the following packages must be imported:
```{r}
library(tidyverse)
library(corrplot)
library(ggplot2)
library(lubridate)
library(gridExtra)
library(caTools)
library(GGally)
```

# Reading the Training and Test Data
From the kaggle competition, we are provided the CSV data for training and testing,
and the dataset is imported like the following:
```{r}
train_data <- read.csv("train.csv")
test_data <- read.csv("test.csv")
```

With the data, we can check if we are missing values, let's first check our training data:
```{r}
# get the missing training values
missing_training_values <- sapply(train_data, function(x) sum(is.na(x)))
# create a data frame with column names and missing values
mtv_df <- data.frame(column = names(missing_training_values), missing_values = missing_training_values)
mtv_df
```
As you can notice there are data in this dataset that contains large amounts of missing values, such
as columns like `PoolQC`, `Fence`, etc. We will be removing the columns with missing values like the following:
```{r}
train_data <- select(train_data, -c(LotFrontage, Alley, MasVnrType, MasVnrArea, BsmtQual , BsmtCond,BsmtExposure, BsmtFinType1, BsmtFinType2, FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageQual, GarageCond, PoolQC, Fence , MiscFeature ))
```

Now it would be good to also check our training data for missing values it may contain:
```{r}
# get the missing values for the testing dataset
missing_testing_values <- sapply(test_data, function (x) sum(is.na(x)))
# create a data frame of the missing testing values
mtestv_df <- data.frame(column = names(missing_testing_values), missing_values = missing_testing_values)
mtestv_df
```

As you can see in our testing dataset we can notice columns that have high amount of missing values such as columns like `LotFrontage`, `Alley`, etc. As before we will be removing the missing values like the following:
```{r}
test_data <- select(test_data, -c(LotFrontage, Alley , MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageQual, GarageCond, PoolQC, Fence, MiscFeature))
```

# Analysis of the Train Data
To analyze the data, we will need to create models using the `lm()` function, where will be using
`SalesPrice` as part of the formula:
```{r}
model <- lm(SalePrice ~., data=train_data)
model
```
